---
title: "eds221-day7-1"
format: html
editor: visual
---

```{r}
library(tidyverse)
library(palmerpenguins)
```

## Data wrangling refresher

1.  Starting from penguins, write a single piped sequence in which you:
    -   Only include penguins at Dream and Biscoe Islands

    -   Remove the year and sex variables

    -   Add a new column body_mass_kg with penguin mass converted from g to kg

    -   Rename the island variable to location

```{r}
new_penguins <- penguins %>% 
  filter(island %in% c("Dream", "Biscoe")) %>% 
  select(-year, -sex) %>% 
  mutate(body_mass_kg = body_mass_g/1000) %>% 
  rename(location = island)
```

2.  Starting from penguins, write a single piped sequence to:
    -   Limit to Adelie penguins

    -   Remove any observations where flipper_length_mm is NA (!is.na() or drop_na)

    -   Group the data by sex

    -   Create a summary table that contains the mean, standard deviation and sample size of flipper lengths for male and female Adelie penguins.

```{r}
penguins_adelie <- penguins %>% 
  filter(species == "Adelie") %>% 
  filter(!is.na(flipper_length_mm)) %>% 
  filter(!is.na(sex)) %>% 
  group_by(species, sex) %>% 
  summarise(mean_flipper_length = mean(flipper_length_mm), sd_flipper_length = sd(flipper_length_mm), count = n())
```

```{r}
animals <- data.frame(
  stringsAsFactors = FALSE,
          location = c("lagoon", "bluff", "creek", "oaks", "bluff"),
           species = c("bobcat", "coyote", "fox", "squirrel", "bobcat"),
          maturity = c("adult", "juvenile", "adult", "juvenile", "adult")
)
```

```{r}
sites <- data.frame(
  stringsAsFactors = FALSE,
          location = c("beach", "lagoon", "bluff", "oaks"),
    full_site_name = c("Goleta Beach","UCSB Lagoon",
                       "Ellwood Mesa","Fremont Campground"),
      jurisdiction = c("SB City", "UCSB", "SB City", "USFS")
)
```

Full join:

uses full_join()

```{r}
full_join_example <- full_join(animals, sites)
```

```{r}
left_join_example <- left_join(animals, sites)
```

```{r}
inner_join_example <- inner_join(animals, sites)
```
